---
title: "Capacity"
output:
  html_document:
    df_print: paged
date created: 12/22/23
created by: Avery Moorhead
last revised by: Erin Mayfield
---

```{r setup, include=FALSE}
###NTS: Currently only running this script on aluminum cement and steel. you can change this by changing rows 222,305, and 341
#Clear objects from workspace
rm(list = ls())

#Set working directory
knitr::opts_knit$set(root.dir = "C:/Users/avery/OneDrive/Desktop/MaterialDemand/Raw Data")

#Load libraries
library(readxl)
library(data.table)
library(dplyr)
library(tidyverse)
library(triangle)
library(stringr)
library(knitr)
```

**Capacity inputs (Bistline et al., 2023):** This code processes electric power capacity input data from Bistline et al. (2023). The data include total capacity and added capacity by technology, capacity expansion model, and year. All capacity data are in units of gigawatts.

```{r message=FALSE, warning=FALSE}

##Cap_IRA: total capacity for each technology, model, and period under IRA scenario (units: gigawatts)

#Load data
Cap_IRA = read_excel("Multi Model Capacity Data.xlsx", 
                      sheet = "Cap IRA", na="ND", range =  cell_rows(1:18), col_names= FALSE)

#Creating dataframe and processing data
Cap_IRA <- t(Cap_IRA)
Cap_IRA <- as.data.frame(Cap_IRA)
rownames(Cap_IRA)<-NULL
new_colnames <- unlist(Cap_IRA[1, ])
colnames(Cap_IRA) <- new_colnames
Cap_IRA <- Cap_IRA[-1,]
colnames(Cap_IRA)[1] ="model"
colnames(Cap_IRA)[2] ="year"
Cap_IRA$year[is.na(Cap_IRA$year)] <- "2021"
Cap_IRA <-  Cap_IRA[(1:34),]
Cap_IRA[is.na(Cap_IRA)] <- 0
Cap_IRA[] <- lapply(Cap_IRA, gsub, pattern = "N/A", replacement = "0", fixed = TRUE)
Cap_IRA <- as.data.frame(Cap_IRA)

#Creating dataframe for total capacity in the year 2021
initial_year <- subset(Cap_IRA, year %in% "2021")
initial_year[is.na(initial_year)] <- '0'


initial_year_melted <- melt(initial_year, id.vars = c("year"), variable.name = "variable", value.name = "value")

initial_year_melted <- initial_year_melted %>%
  mutate(model = NA,  # Add model column with NA
         scenario = NA,  # Add scenario column with NA
         year = 2021) %>%  # Ensure year column is 2021
  select(model, year, variable, value, scenario)  # Select and order columns


initial_year_melted$variable <- gsub("Wind", "onshore wind", initial_year_melted$variable)



#Additional data processing 
Cap_IRA = Cap_IRA[-1,]
Cap_IRA[is.na(Cap_IRA)] <- '0'
Cap_IRA <- rbind(Cap_IRA, initial_year)
Cap_IRA <- reshape2::melt(Cap_IRA, id = c("model","year")) 
Cap_IRA$value <- as.numeric(unlist(Cap_IRA$value))
Cap_IRA$scenario <- "IRA"
Cap_IRA$variable <- gsub("Wind", "onshore wind", Cap_IRA$variable)

##Y_Cap_IRA: additional capacity for each technology, model, and period under IRA scenario (units: gigawatts)

#Load data
Y_Cap_IRA = read_excel("Multi Model Capacity Data.xlsx", 
                     sheet = "Cap IRA", na="ND", col_names= FALSE, range = cell_rows(28:45))

#Creating dataframe and processing data
Y_Cap_IRA <- t(Y_Cap_IRA)
Y_Cap_IRA <- as.data.frame(Y_Cap_IRA)
rownames(Y_Cap_IRA)<-NULL
new_colnames <- unlist(Y_Cap_IRA[1, ])
colnames(Y_Cap_IRA) <- new_colnames
Y_Cap_IRA <- Y_Cap_IRA[-1,]
colnames(Y_Cap_IRA)[1] ="model"
colnames(Y_Cap_IRA)[2] ="year"
Y_Cap_IRA$year[is.na(Y_Cap_IRA$year)] <- "2021"
Y_Cap_IRA <-  Y_Cap_IRA[(1:34),]
Y_Cap_IRA[is.na(Y_Cap_IRA)] <- 0
Y_Cap_IRA[] <- lapply(Y_Cap_IRA, gsub, pattern = "N/A", replacement = "0", fixed = TRUE)
Y_Cap_IRA <- as.data.frame(Y_Cap_IRA)
Y_Cap_IRA[is.na(Y_Cap_IRA)] <- '0'
Y_Cap_IRA <- reshape2::melt(Y_Cap_IRA, id = c("model","year")) 
Y_Cap_IRA$value <- as.numeric(unlist(Y_Cap_IRA$value))
Y_Cap_IRA$scenario <- "IRA"
Y_Cap_IRA$model[Y_Cap_IRA$model=='0']<-"2021"

##Cap_Ref: total capacity for each technology, model, and period under reference scenario (units: gigawatts)
subset_df <- Y_Cap_IRA %>% filter(year == 2025, variable == "Solar Distributed")
average_value <- Y_Cap_IRA %>%
  filter(year == 2025, variable == 'Solar Distributed') %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE))
subset_df <- subset_df %>%
  mutate(average_divided_by_4 = value / 4)


#Load data
Cap_Ref = read_excel("Multi Model Capacity Data.xlsx", 
                     sheet = "Cap Ref", na="ND", range =  cell_rows(1:18), col_names= FALSE)

#Creating dataframe and processing data
Cap_Ref <- t(Cap_Ref)
Cap_Ref <- as.data.frame(Cap_Ref)
rownames(Cap_Ref)<-NULL
new_colnames <- unlist(Cap_Ref[1, ])
colnames(Cap_Ref) <- new_colnames
Cap_Ref <- Cap_Ref[-1,]
colnames(Cap_Ref)[1] ="model"
colnames(Cap_Ref)[2] ="year"
Cap_Ref$year[is.na(Cap_Ref$year)] <- "2021"
Cap_Ref <-  Cap_Ref[(1:34),]
Cap_Ref[is.na(Cap_Ref)] <- 0
Cap_Ref[] <- lapply(Cap_Ref, gsub, pattern = "N/A", replacement = "0", fixed = TRUE)
Cap_Ref <- as.data.frame(Cap_Ref)
Cap_Ref = Cap_Ref[-1,]
Cap_Ref[is.na(Cap_Ref)] <- '0'
Cap_Ref <- rbind(Cap_Ref, initial_year)
Cap_Ref <- reshape2::melt(Cap_Ref, id = c("model","year")) 
Cap_Ref$value <- as.numeric(unlist(Cap_Ref$value))
Cap_Ref$scenario <- "Ref"
Cap_Ref$variable <- gsub("Wind", "onshore wind", Cap_Ref$variable)


##Y_Cap_Ref: additional capacity for each technology, model, and period under reference scenario (units: gigawatts)

#Load data
Y_Cap_Ref = read_excel("Multi Model Capacity Data.xlsx", 
                       sheet = "Cap Ref", na="ND", col_names= FALSE, range = cell_rows(28:45))

#Creating dataframe and processing data
Y_Cap_Ref <- t(Y_Cap_Ref)
Y_Cap_Ref <- as.data.frame(Y_Cap_Ref)
rownames(Y_Cap_Ref)<-NULL
new_colnames <- unlist(Y_Cap_Ref[1, ])
colnames(Y_Cap_Ref) <- new_colnames
Y_Cap_Ref <- Y_Cap_Ref[-1,]
colnames(Y_Cap_Ref)[1] ="model"
colnames(Y_Cap_Ref)[2] ="year"
Y_Cap_Ref$year[is.na(Y_Cap_Ref$year)] <- "2021"
Y_Cap_Ref <-  Y_Cap_Ref[(1:34),]
Y_Cap_Ref[is.na(Y_Cap_Ref)] <- 0
Y_Cap_Ref[] <- lapply(Y_Cap_Ref, gsub, pattern = "N/A", replacement = "0", fixed = TRUE)
Y_Cap_Ref <- as.data.frame(Y_Cap_Ref)
Y_Cap_Ref[is.na(Y_Cap_Ref)] <- '0'
Y_Cap_Ref <- reshape2::melt(Y_Cap_Ref, id = c("model","year")) 
Y_Cap_Ref$value <- as.numeric(unlist(Y_Cap_Ref$value))
Y_Cap_Ref$scenario <- "Ref"
Y_Cap_Ref$model[Y_Cap_Ref$model=='0']<-"2021"

##capacities_all: total capacity for each technology, model, and period under reference and IRA scenarios (units: gigawatts)

#Merging Cap_IRA and Cap_Ref dataframes and data processing
capacities_all <- rbind(Cap_IRA,Cap_Ref)
colnames(capacities_all)[3] ="technology"
capacities_all$technology <- gsub("Wind", "onshore wind", capacities_all$technology)
capacities_all$technology <- gsub("Solar Utility", "utility-scale solar pv", capacities_all$technology)
capacities_all$technology <- gsub("Nuclear Existing", "Nuclear New", capacities_all$technology)

#capacities_all$technology <- gsub("Solar Distributed", "utility-scale solar pv", capacities_all$technology)
#capacities_all$technology <- gsub("NGGT", "Gas", capacities_all$technology)
#capacities_all$technology <- gsub("NGCC", "Gas", capacities_all$technology)
capacities_all$technology <- gsub("NGCC CCS", "Gas CCS", capacities_all$technology)
capacities_all$technology <- gsub("Coal CCS", "Coal CCS", capacities_all$technology)
capacities_all$technology <- gsub("Bio/Other", "Biomass", capacities_all$technology)

#capacities_all$technology <- gsub("Gas CCS", "Gas CCS", capacities_all$technology)

print(unique(capacities_all$technology))

##capacities: mean, minimum, and maximum total capacity across models and for each technology and period under reference and IRA scenarios (units: gigawatts)
capacities <- capacities_all[-1]
capacities <-capacities %>% group_by(scenario,year,technology) %>% summarise(newval = mean(value), max = max(value),min = min(value)) #calculating mean, minimum, and maximum capacities across models 

##capacities_y_all: annualized additional capacity for each technology, model, and period under reference and IRA scenarios (units: gigawatts)

#Merging Y_Cap_IRA and Y_Cap_Ref dataframes
capacities_y_all <- rbind(Y_Cap_IRA,Y_Cap_Ref)

#Annualizing capacity data 
capacities_y_all$newval <- capacities_y_all$value
capacities_y_all$newval[capacities_y_all$year == 2025] <- capacities_y_all$value[capacities_y_all$year == 2025] / 4
capacities_y_all$newval[capacities_y_all$year == 2030] <- capacities_y_all$value[capacities_y_all$year == 2030] / 5
capacities_y_all$newval[capacities_y_all$year == 2035] <- capacities_y_all$value[capacities_y_all$year == 2035] / 5

##capacities_y: mean annualized additional capacity across models and for each technology and period under reference and IRA scenarios (units: gigawatts)
colnames(capacities_y_all)[3] ="technology"
capacities_y_all$technology <- gsub("Wind", "onshore wind", capacities_y_all$technology)


##linear interpolaion for models E4ST and IPM-E
interpolation_df_IRA <- Cap_IRA %>%
  filter(model %in% c("E4ST", "IPM-E"))
interpolation_df_IRA <- rbind(interpolation_df_IRA, initial_year_melted)

interpolation_df_Ref <- Cap_Ref %>%
  filter(model %in% c("E4ST", "IPM-E"))
interpolation_df_Ref <- rbind(interpolation_df_Ref, initial_year_melted)



df_2021 <- interpolation_df_IRA %>% filter(year == 2021)
df_2030 <- interpolation_df_IRA %>% filter(year == 2030)

df_2030_Ref <- interpolation_df_Ref %>% filter(year == 2030)

df_2021$value <- as.numeric(as.character(df_2021$value))
df_2030$value <- as.numeric(as.character(df_2030$value))
df_2030_Ref$value <- as.numeric(as.character(df_2030_Ref$value))

df_2021 <- df_2021 %>% arrange(variable)
df_2030 <- df_2030 %>% arrange(model, variable, scenario)
df_2030_Ref <- df_2030_Ref %>% arrange(model, variable, scenario)


if (any(duplicated(df_2021$variable))) {
  stop("df_2021 contains duplicate technologies.")
}

interpolated_values <- df_2030 %>%
  left_join(df_2021 %>% select(variable, value_2021 = value), by = "variable") %>%
  mutate(value_2025 = value_2021 + (value - value_2021) * (4/9)) %>%
  mutate(delta_2025 = value_2025 - value_2021)

interpolated_values_ref <- df_2030_Ref %>%
  left_join(df_2021 %>% select(variable, value_2021 = value), by = "variable") %>%
  mutate(value_2025 = value_2021 + (value - value_2021) * (4/9)) %>%
  mutate(delta_2025 = value_2025 - value_2021)

new_delta_2030 <- interpolated_values %>%
  mutate(newval = value - value_2025) %>%  # Calculate the new delta
  select(model, year = year, variable, newval, scenario)  # Select and rename columns
new_delta_2030_ref <- interpolated_values_ref %>%
  mutate(newval = value - value_2025) %>%  # Calculate the new delta
  select(model, year = year, variable, newval, scenario)  # Select and rename columns

interpolated_values <- interpolated_values %>%
  select(-value, -value_2021, -value_2025) %>%
  rename(newval = delta_2025) %>% mutate(year = 2025)
final_interpolated_values_IRA <- interpolated_values

interpolated_values_ref <- interpolated_values_ref %>%
  select(-value, -value_2021, -value_2025) %>%
  rename(newval = delta_2025) %>% mutate(year = 2025)
final_interpolated_values_ref <- interpolated_values_ref



final_interpolated_values_IRA <- rbind(new_delta_2030,final_interpolated_values_IRA) %>% rename(technology = variable) %>%  
  mutate(newval = ifelse(year == 2025, newval / 4, ifelse(year == 2030, newval / 5, newval)))

final_interpolated_values_ref <- rbind(new_delta_2030_ref,final_interpolated_values_ref) %>% rename(technology = variable) %>%  
  mutate(newval = ifelse(year == 2025, newval / 4, ifelse(year == 2030, newval / 5, newval)))

capacities_y_all <- capacities_y_all %>% select(-value)
print(colnames(interpolated_values))
capacities_y_all_filtered <- capacities_y_all %>%
  filter(!(year == 2025 | year == 2030) | !(model %in% c("IPM-E", "E4ST")))
capacities_y <- rbind(capacities_y_all_filtered,final_interpolated_values_IRA,final_interpolated_values_ref)
#capacities_y <- capacities_y_all[c(-1,-4)]
capacities_y <-   capacities_y %>% group_by(scenario,year,technology) %>% summarise(newval = mean(newval)) %>% filter(technology != "Wind")




```


```{r}
##multi_model_capacities: additional capacity for each technology, model and year under reference and IRA scenarios (units: gigawatts)

#Creating intermediate dataframes for each period and data processing
temp_df_2025 <- capacities_y %>%
  filter(year == 2025)
duplicated_vector <- rep(1:nrow(temp_df_2025), each = 4)
temp_df_2025 <- temp_df_2025[duplicated_vector, ]

temp_df_2030 <- capacities_y %>%
  filter(year == 2030)
duplicated_vector <- rep(1:nrow(temp_df_2030), each = 5)
temp_df_2030 <- temp_df_2030[duplicated_vector, ]
num_rows <- nrow(temp_df_2030)

temp_df_2035 <- capacities_y %>%
  filter(year == 2035)
duplicated_vector <- rep(1:nrow(temp_df_2035), each = 5)
temp_df_2035 <- temp_df_2035[duplicated_vector, ]
num_rows <- nrow(temp_df_2035)

temp_df_2025 <- temp_df_2025 %>% group_by(technology,scenario)%>%
  mutate(year = ifelse(row_number() %% 3 == 0, 2022, year)) %>% mutate(year = ifelse((row_number() - 1) %% 3 == 0 & row_number() != 1, 2023, year))%>% mutate(year = ifelse((row_number() - 2) %% 3 == 0 & row_number() != 1, 2024, year))

temp_df_2030 <- temp_df_2030 %>% group_by(technology,scenario)%>%
  mutate(year = ifelse(row_number() %% 4 == 0, 2029, year)) %>% mutate(year = ifelse((row_number() - 1) %% 4 == 0 & row_number() != 1, 2028, year))%>% mutate(year = ifelse((row_number() - 2) %% 4 == 0 & row_number() != 1, 2027, year))%>% mutate(year = ifelse((row_number() - 3) %% 4 == 0 & row_number() != 1, 2026, year))

temp_df_2035 <- temp_df_2035 %>% group_by(technology,scenario)%>%
  mutate(year = ifelse(row_number() %% 4 == 0, 2034, year)) %>% mutate(year = ifelse((row_number() - 1) %% 4 == 0 & row_number() != 1, 2033, year))%>% mutate(year = ifelse((row_number() - 2) %% 4 == 0 & row_number() != 1, 2032, year))%>% mutate(year = ifelse((row_number() - 3) %% 4 == 0 & row_number() != 1, 2031, year))

#Combining intermediate dataframes for each period 
multi_model_capacities <- rbind(temp_df_2030,temp_df_2035,temp_df_2025)
multi_model_capacities$technology <- ifelse(multi_model_capacities$technology == "wind", "onshore wind", multi_model_capacities$technology)
multi_model_capacities["technology"][multi_model_capacities["technology"] == "Solar Utility"] <- "utility-scale solar pv"###NTS: check solar distributed vs utility 
#multi_model_capacities["technology"][multi_model_capacities["technology"] == "Solar Distributed"] <- "utility-scale solar pv"###NTS: check solar distributed vs utility 

multi_model_capacities$technology <- gsub("Coal CCS", "Coal CCS", multi_model_capacities$technology)
multi_model_capacities$technology <- gsub("Gas CCS", "Gas CCS", multi_model_capacities$technology)
multi_model_capacities$technology <- gsub("NGCC CCS", "Gas CCS", multi_model_capacities$technology)
#multi_model_capacities$technology <- gsub("NGCC", "Gas", multi_model_capacities$technology)
multi_model_capacities$technology <- gsub("Bio/Other", "Biomass", multi_model_capacities$technology)

multi_model_capacities <- multi_model_capacities %>%
  group_by(year, scenario, technology) %>%
  summarise(newval = sum(newval))

print(unique(multi_model_capacities$technology))
```

**Capacity inputs (Jenkins et al., 2023):** This code uploads and processes electric power capacity input data from Jenkins et al. (2023). The data include total capacity and added capacity by technology, capacity expansion model, and year. All capacity data are in units of gigawatts.

```{r message=FALSE, warning=FALSE}
##repeat_capacities: total capacity for each technology and year under reference and IRA scenarios (units: gigawatts)

#Load data
input_capacities= read.csv("capacity_y.csv")

#Creating dataframe and processing data
scenarios = c("Ref", "IRA")#Define the REPEAT scenarios to be included
repeat_technologies = c("onshore wind", "utility-scale solar pv", "Nuclear New", "Wind Offshore", "Gas","Coal","offshore wind")#Define the REPEAT technologies to be included
repeat_scenarios = c("baseline", "ira_mid")
input_capacities["tech..outputs_group_detailed"][input_capacities["tech..outputs_group_detailed"] == "offshore wind"] <- "offshore wind fixed"
input_capacities["tech..outputs_group_detailed"][input_capacities["tech..outputs_group_detailed"] == "new nuclear"] <- "Nuclear New"
input_capacities$tech..outputs_group_detailed <- gsub("existing_gas combustion turbine", "Gas", input_capacities$tech..outputs_group_detailed)
input_capacities$tech..outputs_group_detailed <- gsub("existing_gas combined cycle", "Gas", input_capacities$tech..outputs_group_detailed)
input_capacities$tech..outputs_group_detailed <- gsub("existing_gas steam turbine", "Gas", input_capacities$tech..outputs_group_detailed)
#input_capacities$tech..outputs_group_detailed <- gsub("gas production", "Gas", input_capacities$tech..outputs_group_detailed)
input_capacities$tech..outputs_group_detailed <- gsub("naturalgas_ccavgcf", "Gas", input_capacities$tech..outputs_group_detailed)
input_capacities$tech..outputs_group_detailed <- gsub("naturalgas_ctavgcf", "Gas", input_capacities$tech..outputs_group_detailed)
input_capacities$tech..outputs_group_detailed <- gsub("offshore wind fixed", "offshore wind", input_capacities$tech..outputs_group_detailed)
input_capacities$tech..outputs_group_detailed <- gsub("offshore wind floating", "offshore wind", input_capacities$tech..outputs_group_detailed)
input_capacities$tech..outputs_group_detailed <- gsub("rooftop solar", "utility-scale solar pv", input_capacities$tech..outputs_group_detailed)




print(unique(input_capacities$tech..outputs_group_detailed))


repeat_capacities = subset(input_capacities, run.name %in% repeat_scenarios & tech..outputs_group_detailed %in% repeat_technologies) %>% group_by(run.name,year,tech..outputs_group_detailed) %>% summarise(newval = sum(value))

##repeat_capacities_new: annualized added capacity for each technology and year under reference and IRA scenarios (units: gigawatts)

#Load data
input_capacities_delta = read.csv("capacity_delta_y.csv")

#Creating dataframe and processing data
input_capacities_delta["tech..outputs_group_detailed"][input_capacities_delta["tech..outputs_group_detailed"] == "offshore wind"] <- "offshore wind fixed"
input_capacities_delta["tech..outputs_group_detailed"][input_capacities_delta["tech..outputs_group_detailed"] == "new nuclear"] <- "Nuclear New"
#input_capacities_delta$tech..outputs_group_detailed <- gsub("gas production", "Gas", input_capacities_delta$tech..outputs_group_detailed)
input_capacities_delta$tech..outputs_group_detailed <- gsub("naturalgas_ccavgcf", "Gas", input_capacities_delta$tech..outputs_group_detailed)
input_capacities_delta$tech..outputs_group_detailed <- gsub("naturalgas_ctavgcf", "Gas", input_capacities_delta$tech..outputs_group_detailed)
input_capacities_delta$tech..outputs_group_detailed <- gsub("offshore wind fixed", "offshore wind", input_capacities_delta$tech..outputs_group_detailed)
input_capacities_delta$tech..outputs_group_detailed <- gsub("offshore wind floating", "offshore wind", input_capacities_delta$tech..outputs_group_detailed)
print(unique(input_capacities_delta$tech..outputs_group_detailed))
#input_capacities_delta$tech..outputs_group_detailed <- gsub("lng production", "Gas", input_capacities_delta$tech..outputs_group_detailed)
input_capacities_delta$tech..outputs_group_detailed <- gsub("existing_gas combined cycle", "Gas", input_capacities_delta$tech..outputs_group_detailed)
input_capacities_delta$tech..outputs_group_detailed <- gsub("existing_gas combustion turbine", "Gas", input_capacities_delta$tech..outputs_group_detailed)
input_capacities_delta$tech..outputs_group_detailed <- gsub("existing_gas steam turbine", "Gas", input_capacities_delta$tech..outputs_group_detailed)
input_capacities_delta$tech..outputs_group_detailed <- gsub("rooftop solar", "utility-scale solar pv", input_capacities_delta$tech..outputs_group_detailed)





repeat_capacities_new <- subset(input_capacities_delta, run.name %in% repeat_scenarios & type == "new" & tech..outputs_group_detailed %in% repeat_technologies) %>%
  group_by(run.name, year, tech..outputs_group_detailed) %>%
  summarise(newval = sum(value))

repeat_capacities_new <- rename(repeat_capacities_new, technology = tech..outputs_group_detailed)
repeat_capacities_new <- filter(repeat_capacities_new, year %in% c("2026", "2030", "2035"))
repeat_capacities_new <- mutate(repeat_capacities_new, newval = case_when(
  year == 2026 ~ newval / 2,
  year == 2030 ~ newval / 2,
  year == 2035 ~ newval / 3
))
repeat_capacities_new <- mutate(repeat_capacities_new, year = ifelse(year == 2026, 2025, year))

repeat_capacities_new <- group_by(repeat_capacities_new, year, run.name, technology) %>%
  summarise(newval = sum(newval), .groups = 'drop')
new_rows <- data.frame(
  year = c(2025, 2025),
  run.name = c('ira_mid', 'baseline'),
  technology = c('Nuclear New', 'Nuclear New'),
  newval = c(0, 0)
)

print(unique(input_capacities_delta$tech..outputs_group_detailed))


```
```{r}
#adding initial year to multi model df
melted_initial_year <- melt(initial_year, id.vars = c("year"), variable.name = "technology")
melted_initial_year <- melted_initial_year[-1,]

names(melted_initial_year)[names(melted_initial_year) == "value"] <- "newval"
melted_initial_year <- melted_initial_year %>%
                         mutate(scenario = NA)
summary(melted_initial_year)
melted_initial_year$newval <- as.numeric(melted_initial_year$newval)

melted_initial_year$technology <- gsub("Nuclear\\.Existing", "Nuclear Existing", melted_initial_year$technology)
melted_initial_year$technology <- gsub("Nuclear\\.New", "Nuclear New", melted_initial_year$technology)
melted_initial_year$technology <- gsub("NGCC.CCS", " Gas CCS", melted_initial_year$technology)
melted_initial_year$technology <- gsub("Hydrogen.Green", "Hydrogen Green", melted_initial_year$technology)
melted_initial_year$technology <- gsub("Solar.Utility", "utility-scale solar pv", melted_initial_year$technology)
melted_initial_year$technology <- gsub("Wind", "onshore wind", melted_initial_year$technology)



# Print the updated data frame
print(melted_initial_year)


print(melted_initial_year)
print(multi_model_capacities)
#combined_df <- rbind(melted_initial_year, multi_model_capacities)
#multi_model_capacities<-combined_df

```

**Saving data frames to "inputs" folder:** This code saves the dataframes for use in the model 

```{r}
folder_path <- "C:/Users/avery/OneDrive/Desktop/MaterialDemand/Inputs"

file_path1 <- file.path(folder_path, "initial_year.csv")
file_path2 <- file.path(folder_path, "multi_model_capacities.csv")
file_path3 <- file.path(folder_path, "repeat_capacities_new.csv")


# Write each dataframe to its respective CSV file in the specified folder
write.csv(initial_year, file = file_path1, row.names = FALSE)
write.csv(multi_model_capacities, file = file_path2, row.names = FALSE)
write.csv(repeat_capacities_new, file = file_path3, row.names = FALSE)

```
```{r}



```

