---
title: "Material Intensity"
output:
  html_document:
    df_print: paged
date created: 12/22/23
created by: Avery Moorhead
last revised by: Erin Mayfield
---
```{r setup, include=FALSE}
library(here)
library(readxl)
library(data.table)
library(dplyr)
library(tidyverse)
library(stringr)
library(knitr)

knitr::opts_knit$set(root.dir = "C:/Users/avery/OneDrive/Desktop/MaterialDemand/Raw Data")

# Clear objects from workspace
rm(list = ls())

```


**Historical material supply, demand, imports, and exports:** This code uploads historical U.S. consumption, production, imports, and exports of materials for 2018 to 2022 from the U.S. Geological Survey (2023).

```{r message=FALSE, warning=FALSE}
##inputs_usgs: U.S. supply, demand, imports, and exports for materials from 2018 to 2021

#Load data

input_usgs = read_excel("US Geological Survey 2023/Data by material/USGS_2023_IRA materials calculations.xlsx", 
                        sheet = "Combined", range = cell_cols(1:5))
#Filtering of USGS data
usgs_variables = c("Consumption - US", "Production - US", "Import - US", "Export - US")
flow = subset(input_usgs, Variable %in%usgs_variables)
```

**Material intensity (Liang et al., 202X; Wang et al., 202X):** This code uploads and processes material intensity values from Liang et al. (202X) and Wang et al. (202X).

```{r message=FALSE, warning=FALSE}
##repeat_capacities_delta: annualized added capacity for each technology and year under reference and IRA scenarios (units: gigawatts)

#Load data
material_intensity = read_excel("Final Intensity Values.xlsx", sheet = "Intensity Values", range = cell_cols(1:6))
unique(material_intensity$technology)
#Creating dataframe and processing data
colnames(material_intensity)[1] ="technology"
colnames(material_intensity)[3] ="value"
material_intensity <- material_intensity[!duplicated(material_intensity), ]
material_intensity <- material_intensity[material_intensity$`Year of the data` <= 2025, 
                                  !(names(material_intensity) %in% c('Year of the data', 'Authors', 'Publish year'))]
###NTS:data processing, worth checking some of these 
material_intensity["technology"][material_intensity["technology"] == "Nuclear"] <- "Nuclear New"
material_intensity["technology"][material_intensity["technology"] == "Onshore_AG"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Onshore_DD_PMG"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Wind Directed_Drive"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Geared"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Wind onshore"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Wind Onshore"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Onshore Wind"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Offshore Wind"] <- "offshore wind"
material_intensity["technology"][material_intensity["technology"] == "CoalCCS"] <- "Coal CCS"
material_intensity["technology"][material_intensity["technology"] == "GasCCS"] <- "Gas CCS"
material_intensity["technology"][material_intensity["technology"] == "Gas"] <- "NGCC"


material_intensity["technology"][material_intensity["technology"] == "CSI_PV"] <- "utility-scale solar pv"
material_intensity["technology"][material_intensity["technology"] == "BiomassCCS"] <- "Bio CCS"

material_intensity$technology[material_intensity$technology == "Offshore_DD_PMG"] <- "offshore wind"
material_intensity$technology[material_intensity$technology == "Offshore_AG"] <- "offshore wind"
material_intensity$Material[material_intensity$Material == "Neodymium "] <-  "Neodymium"
material_intensity$Material <- gsub("\\s+", "", material_intensity$Material)
material_intensity$Material <- gsub("[^[:alnum:]]", "", material_intensity$Material)

print(unique(material_intensity$technology))

solar_pv_rows <- material_intensity %>%
  filter(technology == 'utility-scale solar pv') %>%
  mutate(technology = 'Solar Distributed')


NGCC_rows <- material_intensity %>%
  filter(technology == 'NGCC') %>%
  mutate(technology = 'NGGT')
Gas_rows <- material_intensity %>%
  filter(technology == 'NGCC') %>%
  mutate(technology = 'Gas')





material_intensity <- bind_rows(material_intensity, solar_pv_rows)
material_intensity <- bind_rows(material_intensity, NGCC_rows)
material_intensity <- bind_rows(material_intensity, Gas_rows)

print(unique(material_intensity$technology))

```
```{r}
library(fitdistrplus)
test = subset(material_intensity, technology == "utility-scale solar pv" & Material == "Silver")
fw <- fitdist(test$value, "weibull", "mge")
summary(fw)
```

```{r}
library(dplyr)

# Sensitivity analysis plots
par(mfrow = c(3, 3))
par(mar = c(2, 2, 2, 2))

# Filter dataframe to only include specified technologies
filtered_df <- filter(material_intensity, technology %in% c("onshore wind", "offshore wind", "utility-scale solar pv", "nuclear"))

for (tech in unique(filtered_df$technology)) {
  for (material in unique(filtered_df$Material)) {
    tryCatch(
      expr = {
        combo_type <- paste(material, tech, sep = ".")
        combo_type <- gsub(" ", ".", combo_type)
        combo_type <- gsub("-", ".", combo_type)

        hist_type <- gsub("-", ".", combo_type)
        hist_type <- gsub(" ", ".", hist_type)

        axis_labels <- subset(filtered_df, technology == tech & Material == material)
        hist_df <- subset(filtered_df, technology == tech & Material == material, select = c("value"))

        hist(hist_df$value, col = "antiquewhite1", cex.main = 1.4, xaxt = 'n', xlab = '', main = combo_type, xlim = c(min(axis_labels$value), max(axis_labels$value)))
        axis(1, at = c(mean(axis_labels$value), min(axis_labels$value), max(axis_labels$value)),
             labels = c(round(mean(axis_labels$value), digits = 5), round(min(axis_labels$value), digits = 5), round(max(axis_labels$value), digits = 5)))
      },
      error = function(e) {
        dummy <- '1'
      }
    )
  }
}


```


**Saving data frames to "inputs" folder:** This code saves the dataframes for use in the model 
```{r}
folder_path <- "C:/Users/avery/OneDrive/Desktop/MaterialDemand/Inputs"
file_path1 <- file.path(folder_path, "input_usgs.csv")
file_path2 <- file.path(folder_path, "intensity_data.csv")

write.csv(input_usgs, file = file_path1, row.names = FALSE)
write.csv(material_intensity, file = file_path2, row.names = FALSE)

```


```{r}

library(dplyr)
library(fitdistrplus)


# Group by technology and Material and count the number of entries
count_table <- material_intensity %>%
  group_by(technology, Material) %>%
  summarise(count = n())
material_intensity_dist_df <- material_intensity %>%
  inner_join(count_table, by = c("technology", "Material")) %>%
  filter(count >= 30)

solarsilver <- material_intensity_dist_df %>%
  filter(technology == "onshore wind" & Material == "Steel")

descdist(solarsilver$value, boot = 1000)

fw <- fitdist(solarsilver$value, "weibull")
summary(fw)

par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))
#fg <- fitdist(solarsilver$value, "gamma")
fln <- fitdist(solarsilver$value, "lnorm")
plot.legend <- c("Weibull", "lognormal", "gamma")
denscomp(list(fw, fln, fg), legendtext = plot.legend)
qqcomp(list(fw, fln, fg), legendtext = plot.legend)
cdfcomp(list(fw, fln, fg), legendtext = plot.legend)
ppcomp(list(fw, fln, fg), legendtext = plot.legend)





```

```{r}
#solar and silver
count_table <- material_intensity %>%
  group_by(technology, Material) %>%
  summarise(count = n())
material_intensity_dist_df <- material_intensity %>%
  inner_join(count_table, by = c("technology", "Material")) %>%
  subset(count >= 30)

solarsilver <- subset(material_intensity, technology == "utility-scale solar pv" & Material == "Silver")

#descdist(solarsilver$value, boot = 1000)

#fitting distributions
fw <- fitdist(solarsilver$value, "weibull")
summary(fw)
#fg <- fitdist(solarsilver$value, "gamma")
#fln <- fitdist(solarsilver$value, "lnorm")
#fe <- fitdist(solarsilver$value, "exp")

#plotting
#plot.legend <- c("Weibull", "lognormal", "gamma", "exponential")
#denscomp(list(fw, fln, fg), legendtext = plot.legend)
#qqcomp(list(fw, fln, fg,fe), legendtext = plot.legend)
#cdfcomp(list(fw, fln, fg), legendtext = plot.legend)
#ppcomp(list(fw, fln, fg,fe), legendtext = plot.legend)


```

```{r}

```

```{r}
#onshore wind and copper
onshorewindcopper <- material_intensity_dist_df %>%
  filter(technology == "onshore wind" & Material == "Copper")
library(fitdistrplus)
descdist(onshorewindcopper$value, boot = 1000)

#fitting distributions
fw <- fitdist(onshorewindcopper$value, "weibull")
#fg <- fitdist(onshorewindcopper$value, "gamma")gamma doesn't work 
fln <- fitdist(onshorewindcopper$value, "lnorm")
#fe <- fitdist(onshorewindcopper$value, "exp") exp doesn't work either

#plotting
plot.legend <- c("Weibull", "lognormal", "gamma")
#denscomp(list(fw, fln, fg), legendtext = plot.legend)
qqcomp(list(fw, fln), legendtext = plot.legend)
#cdfcomp(list(fw, fln, fg), legendtext = plot.legend)
ppcomp(list(fw, fln), legendtext = plot.legend)


```
```{r}
#offshore wind and copper

dist_df <- material_intensity_dist_df %>%
  filter(technology == "offshore wind" & Material == "Copper")
library(fitdistrplus)
descdist(dist_df$value, boot = 1000)
#fitting distributions
fw <- fitdist(dist_df$value, "weibull")
#fe <- fitdist(dist_df$value, "exp")
#fg <- fitdist(dist_df$value, "gamma") #gamma doesn't work 
fln <- fitdist(dist_df$value, "lnorm")

#plotting
plot.legend <- c("Weibull", "lognormal")
#denscomp(list(fw, fln, fg), legendtext = plot.legend)
qqcomp(list(fw, fln), legendtext = plot.legend)
#cdfcomp(list(fw, fln, fg), legendtext = plot.legend)
ppcomp(list(fw, fln), legendtext = plot.legend)

```
```{r}
#onshore wind neodymium

dist_df <- material_intensity_dist_df %>%
  filter(technology == "offshore wind" & Material == "Neodymium")
library(fitdistrplus)
#descdist(dist_df$value, boot = 1000)

#fitting distributions
fw <- fitdist(dist_df$value, "weibull")
fg <- fitdist(dist_df$value, "gamma") #gamma doesn't work 
fln <- fitdist(dist_df$value, "lnorm")
fn <- fitdist(dist_df$value, "norm")
fe <- fitdist(dist_df$value, "exp")

#plotting
plot.legend <- c("Weibull","Gamma", "Lognormal","Normal","Exponential")
denscomp(list(fw,fg, fln, fn,fe), main = "Neodymium, Offshore Wind", legendtext= NULL)

```
```{r}

```

```{r}
#onshore wind steel

dist_df <- material_intensity_dist_df %>%
  filter(technology == "onshore wind" & Material == "Steel")
library(fitdistrplus)
descdist(dist_df$value, boot = 1000)

#fitting distributions
fw <- fitdist(dist_df$value, "weibull")
#fe <- fitdist(dist_df$value, "exp")
#fg <- fitdist(dist_df$value, "gamma") #gamma doesn't work 
fln <- fitdist(dist_df$value, "lnorm")

#plotting
plot.legend <- c("Weibull", "lognormal")
#denscomp(list(fw, fln, fg), legendtext = plot.legend)
qqcomp(list(fw, fln), legendtext = plot.legend)
#cdfcomp(list(fw, fln, fg), legendtext = plot.legend)
ppcomp(list(fw, fln), legendtext = plot.legend)

```

```{r}
#onshore wind dysprosium 

dist_df <- material_intensity_dist_df %>%
  filter(technology == "onshore wind" & Material == "Dysprosium")
library(fitdistrplus)
descdist(dist_df$value, boot = 1000)

#fitting distributions
fw <- fitdist(dist_df$value, "weibull")
fe <- fitdist(dist_df$value, "exp")
fg <- fitdist(dist_df$value, "gamma") 
fln <- fitdist(dist_df$value, "lnorm")

#plotting
plot.legend <- c("Weibull", "lognormal", "gamma","exponential")
denscomp(list(fw, fln, fg), legendtext = plot.legend)
qqcomp(list(fw, fln,fg,fe), legendtext = plot.legend)
cdfcomp(list(fw, fln, fg), legendtext = plot.legend)
ppcomp(list(fw, fln,fg,fe), legendtext = plot.legend)

```
```{r}
#offshore wind neodymium

dist_df <- material_intensity_dist_df %>%
  filter(technology == "Wind Offshore" & Material == "Neodymium")
library(fitdistrplus)
descdist(dist_df$value, boot = 1000)

#fitting distributions
fw <- fitdist(dist_df$value, "weibull")
fe <- fitdist(dist_df$value, "exp")
fg <- fitdist(dist_df$value, "gamma") 
fln <- fitdist(dist_df$value, "lnorm")

#plotting
plot.legend <- c("Weibull", "lognormal", "gamma","exponential")
denscomp(list(fw, fln, fg,fe), legendtext = plot.legend)
qqcomp(list(fw, fln,fg,fe), legendtext = plot.legend)
cdfcomp(list(fw, fln, fg,fe), legendtext = plot.legend)
ppcomp(list(fw, fln,fg,fe), legendtext = plot.legend)

```
```{r}
#onshore wind aluminum 

dist_df <- material_intensity_dist_df %>%
  filter(technology == "onshore wind" & Material == "Aluminum")
library(fitdistrplus)
descdist(dist_df$value, boot = 1000)

#fitting distributions
fw <- fitdist(dist_df$value, "weibull")
#fe <- fitdist(dist_df$value, "exp")
#fg <- fitdist(dist_df$value, "gamma") #gamma doesn't work 
fln <- fitdist(dist_df$value, "lnorm")

#plotting
plot.legend <- c("Weibull", "lognormal")
#denscomp(list(fw, fln, fg), legendtext = plot.legend)
qqcomp(list(fw, fln), legendtext = plot.legend)
#cdfcomp(list(fw, fln, fg), legendtext = plot.legend)
ppcomp(list(fw, fln), legendtext = plot.legend)

```

```{r}

#data processing for under 30 technology/material combinations
library(dplyr)

subset_material_intensity <- material_intensity %>%
  filter(technology %in% c("Wind Offshore", "Onshore Wind", "Utility-Scale Solar PV", "Nuclear New"))

count_table <- subset_material_intensity %>%
  group_by(technology, Material) %>%
  tally()

under30_material_intensity <- subset_material_intensity %>%
  inner_join(count_table, by = c("technology", "Material")) %>%
  filter(n <= 30) 

# Now material_intensity_dist_df contains the subsetted data with combinations having 30 or fewer occurrences for the specified technologies

```

```{r}
#test version of bootstrapping 
under30_dist_df <- under30_material_intensity %>%
  filter(technology == "Wind Offshore" & Material == "Dysprosium")
#fendo.ln <- fitdist(under30_dist_df$value, "lnorm")
bootstrappedvalues <- sample(under30_dist_df$value, 100, replace = TRUE)

fendo.ln <- fitdist(bootstrappedvalues, "lnorm", lower = c(0, 1))

#bootstrap_dist_df <- bootdist(fendo.ln, niter = 1001)
#summary(bootstrap_dist_df)



```



```

