---
title: "Material Intensity"
output:
  html_document:
    df_print: paged
date created: 12/22/23
created by: Avery Moorhead
last revised by: Erin Mayfield
---
```{r setup, include=FALSE}
library(here)
library(readxl)
library(data.table)
library(dplyr)
library(tidyverse)
library(stringr)
library(knitr)
library(fitdistrplus)

knitr::opts_knit$set(root.dir = "C:/Users/avery/OneDrive/Desktop/MaterialDemand/Raw Data")

# Clear objects from workspace
rm(list = ls())

```


**Historical material supply, demand, imports, and exports:** This code uploads historical U.S. consumption, production, imports, and exports of materials for 2018 to 2022 from the U.S. Geological Survey (2023).

```{r message=FALSE, warning=FALSE}
##inputs_usgs: U.S. supply, demand, imports, and exports for materials from 2018 to 2021

#Load data

input_usgs = read_excel("US Geological Survey 2023/Data by material/USGS_2023_IRA materials calculations.xlsx", 
                        sheet = "Combined", range = cell_cols(1:5))
#Filtering of USGS data
usgs_variables = c("Consumption - US", "Production - US", "Import - US", "Export - US")
flow = subset(input_usgs, Variable %in%usgs_variables)
```

**Material intensity (Liang et al., 202X; Wang et al., 202X):** This code uploads and processes material intensity values from Liang et al. (202X) and Wang et al. (202X).

```{r message=FALSE, warning=FALSE}
#Load data
material_intensity = read_excel("Final Intensity Values.xlsx", sheet = "Intensity Values", range = cell_cols(1:6))
unique(material_intensity$technology)

#Creating dataframe and processing data
colnames(material_intensity)[1] ="technology"
colnames(material_intensity)[3] ="value"
material_intensity <- material_intensity[!duplicated(material_intensity), ]
material_intensity <- material_intensity[material_intensity$`Year of the data` <= 2025, 
                                  !(names(material_intensity) %in% c('Year of the data', 'Authors', 'Publish year'))]
material_intensity["technology"][material_intensity["technology"] == "Nuclear"] <- "Nuclear New"
material_intensity["technology"][material_intensity["technology"] == "Onshore_AG"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Onshore_DD_PMG"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Wind Directed_Drive"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Geared"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Wind onshore"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Wind Onshore"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Onshore Wind"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Offshore Wind"] <- "offshore wind"
material_intensity["technology"][material_intensity["technology"] == "CoalCCS"] <- "Coal CCS"
material_intensity["technology"][material_intensity["technology"] == "GasCCS"] <- "Gas CCS"
material_intensity["technology"][material_intensity["technology"] == "Gas"] <- "NGCC"
material_intensity["technology"][material_intensity["technology"] == "CSI_PV"] <- "utility-scale solar pv"
material_intensity["technology"][material_intensity["technology"] == "BiomassCCS"] <- "Bio CCS"
material_intensity$technology[material_intensity$technology == "Offshore_DD_PMG"] <- "offshore wind"
material_intensity$technology[material_intensity$technology == "Offshore_AG"] <- "offshore wind"
material_intensity$Material[material_intensity$Material == "Neodymium "] <-  "Neodymium"
material_intensity$Material <- gsub("\\s+", "", material_intensity$Material)
material_intensity$Material <- gsub("[^[:alnum:]]", "", material_intensity$Material)

#mutating rows for additional technologies 
solar_pv_rows <- material_intensity %>%
  filter(technology == 'utility-scale solar pv') %>%
  mutate(technology = 'Solar Distributed')

NGCC_rows <- material_intensity %>%
  filter(technology == 'NGCC') %>%
  mutate(technology = 'NGGT')
Gas_rows <- material_intensity %>%
  filter(technology == 'NGCC') %>%
  mutate(technology = 'Gas')

material_intensity <- bind_rows(material_intensity, solar_pv_rows)
material_intensity <- bind_rows(material_intensity, NGCC_rows)
material_intensity <- bind_rows(material_intensity, Gas_rows)
```
**Saving data frames to "inputs" folder:** This code saves the dataframes for use in the model 
```{r}
folder_path <- "C:/Users/avery/OneDrive/Desktop/MaterialDemand/Inputs"
file_path1 <- file.path(folder_path, "input_usgs.csv")
file_path2 <- file.path(folder_path, "intensity_data.csv")

write.csv(input_usgs, file = file_path1, row.names = FALSE)
write.csv(material_intensity, file = file_path2, row.names = FALSE)

```

**Material intensity distribution fitting plots code setup**
```{r}
count_table <- material_intensity %>%
  group_by(technology, Material) %>%
  summarise(count = n())
material_intensity_dist_df <- material_intensity %>%
  inner_join(count_table, by = c("technology", "Material")) %>%
  filter(count >= 30)

```

*Distribution Fitting Plots*
```{r}

dist_df <- material_intensity_dist_df %>%
  filter(technology == "offshore wind" & Material == "Neodymium")
library(fitdistrplus)

#fitting distributions
fw <- fitdist(dist_df$value, "weibull")
fg <- fitdist(dist_df$value, "gamma") #gamma doesn't work 
fln <- fitdist(dist_df$value, "lnorm")
fn <- fitdist(dist_df$value, "norm")
fe <- fitdist(dist_df$value, "exp")

#plotting
plot.legend <- c("Weibull","Gamma", "Lognormal","Normal","Exponential")
denscomp(list(fw,fg, fln, fn,fe), main = "Neodymium, Offshore Wind", legendtext= plot.legend)

```




