---
title: "Material Intensity"
output:
  html_document:
    df_print: paged
date created: 12/22/23
created by: Avery Moorhead
last revised by: Erin Mayfield
---
```{r setup, include=FALSE}
# Load necessary packages
library(here)
library(readxl)
library(data.table)
library(dplyr)
library(tidyverse)
library(stringr)
library(knitr)

# Set working directory using here::set_here()
knitr::opts_knit$set(root.dir = "C:/Users/avery/OneDrive/Desktop/MaterialDemand/Raw Data")

# Clear objects from workspace
rm(list = ls())

```


**Historical material supply, demand, imports, and exports:** This code uploads historical U.S. consumption, production, imports, and exports of materials for 2018 to 2022 from the U.S. Geological Survey (2023).

```{r message=FALSE, warning=FALSE}
##inputs_usgs: U.S. supply, demand, imports, and exports for materials from 2018 to 2021

#Load data

input_usgs = read_excel("US Geological Survey 2023/Data by material/USGS_2023_IRA materials calculations.xlsx", 
                        sheet = "Combined", range = cell_cols(1:5))
#Filtering of USGS data
usgs_variables = c("Consumption - US", "Production - US", "Import - US", "Export - US")
flow = subset(input_usgs, Variable %in%usgs_variables)
```

**Material intensity (Liang et al., 202X; Wang et al., 202X):** This code uploads and processes material intensity values from Liang et al. (202X) and Wang et al. (202X).

```{r message=FALSE, warning=FALSE}
##repeat_capacities_delta: annualized added capacity for each technology and year under reference and IRA scenarios (units: gigawatts)

#Load data
material_intensity = read_excel("Final Intensity Values.xlsx", sheet = "Intensity Values", range = cell_cols(1:6))

#Creating dataframe and processing data
colnames(material_intensity)[1] ="technology"
colnames(material_intensity)[3] ="value"
material_intensity <- material_intensity[!duplicated(material_intensity), ]
material_intensity <- material_intensity[material_intensity$`Year of the data` <= 2025, 
                                  !(names(material_intensity) %in% c('Year of the data', 'Authors', 'Publish year'))]
###NTS:data processing, worth checking some of these 
material_intensity["technology"][material_intensity["technology"] == "Nuclear"] <- "Nuclear New"
material_intensity["technology"][material_intensity["technology"] == "Onshore_AG"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Onshore_DD_PMG"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Wind Directed_Drive"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Geared"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Wind onshore"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Wind Onshore"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "Onshore Wind"] <- "onshore wind"
material_intensity["technology"][material_intensity["technology"] == "CSI_PV"] <- "utility-scale solar pv"
material_intensity$technology[material_intensity$technology == "Offshore_DD_PMG"] <- "Wind Offshore"
material_intensity$technology[material_intensity$technology == "Offshore_AG"] <- "Wind Offshore"
material_intensity$Material[material_intensity$Material == "Neodymium "] <-  "Neodymium"

library(dplyr)

# Group by technology and Material and count the number of entries
count_table <- material_intensity %>%
  group_by(technology, Material) %>%
  summarise(count = n())

# Print the new table
print(count_table)

```
```{r}
# Load necessary library
library(dplyr)

# Sensitivity analysis plots
par(mfrow = c(3, 3))
par(mar = c(2, 2, 2, 2))

# Filter dataframe to only include specified technologies
filtered_df <- filter(material_intensity, technology %in% c("onshore wind", "Offshore Wind", "utility-scale solar pv", "nuclear"))

for (tech in unique(filtered_df$technology)) {
  for (material in unique(filtered_df$Material)) {
    tryCatch(
      expr = {
        combo_type <- paste(material, tech, sep = ".")
        combo_type <- gsub(" ", ".", combo_type)
        combo_type <- gsub("-", ".", combo_type)

        hist_type <- gsub("-", ".", combo_type)
        hist_type <- gsub(" ", ".", hist_type)

        axis_labels <- subset(filtered_df, technology == tech & Material == material)
        hist_df <- subset(filtered_df, technology == tech & Material == material, select = c("value"))

        hist(hist_df$value, col = "antiquewhite1", cex.main = 1.4, xaxt = 'n', xlab = '', main = combo_type, xlim = c(min(axis_labels$value), max(axis_labels$value)))
        axis(1, at = c(mean(axis_labels$value), min(axis_labels$value), max(axis_labels$value)),
             labels = c(round(mean(axis_labels$value), digits = 5), round(min(axis_labels$value), digits = 5), round(max(axis_labels$value), digits = 5)))
      },
      error = function(e) {
        dummy <- '1'
      }
    )
  }
}


```


**Saving data frames to "inputs" folder:** This code saves the dataframes for use in the model 
```{r}
folder_path <- "C:/Users/avery/OneDrive/Desktop/MaterialDemand/Inputs"
file_path1 <- file.path(folder_path, "input_usgs.csv")
file_path2 <- file.path(folder_path, "intensity_data.csv")

# Write each dataframe to its respective CSV file in the specified folder
write.csv(input_usgs, file = file_path1, row.names = FALSE)
write.csv(material_intensity, file = file_path2, row.names = FALSE)

```
```{r}

```

