---
title: "Material Demand Calculations"
author: "Avery Moorhead"
date: "`12/22/23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/avery/OneDrive/Desktop/MaterialDemand/Inputs")

###NTS: Currently only running this script on aluminum cement and steel. you can change this by changing rows 222,305, and 341
#Clear objects from workspace
rm(list = ls())

#Load libraries
library(readxl)
library(data.table)
library(dplyr)
library(tidyverse)
library(triangle)
library(stringr)
library(knitr)

```

**Setup** Reading data Frames from "Capacities" and "Material Intensity. Defining technologies and scenarios


```{r message=FALSE, warning=FALSE}
# Print the current working directory
print(getwd())
input_usgs<- read.csv("input_usgs.csv")
material_intensity<- read.csv("intensity_data.csv")
initial_year<- read.csv("initial_year.csv")
multi_model_capacities<- read.csv("multi_model_capacities.csv")
repeat_capacities_new<- read.csv("repeat_capacities_new.csv")



scenarios = c("Ref", "IRA")#Define the REPEAT scenarios to be included
repeat_technologies = c("onshore wind", "utility-scale solar pv", "Nuclear New", "Wind Offshore")#Define the REPEAT technologies to be included
repeat_scenarios = c("baseline", "ira_mid")
usgs_variables = c("Consumption - US", "Production - US", "Import - US", "Export - US")


```

**Material intensity summary statistics and distribution fitting:** This code estimates summary statistics and fits probability distributions to material intensity data from Liang et al. (202X) and Wang et al. (202X).

```{r message=FALSE, warning=FALSE}
#Subset material intensity data to include some materials 
materials <- c("Aluminum","Cement","Steel")
#print(materials)# Print the unique materials
intensity = subset(material_intensity, Material %in% materials)

#Check if all material/technology scenarios are present, if not add 0.
for (material in materials) {
  for (tech in repeat_technologies){  
    check_empty_frame = subset(intensity, technology %in% tech & Material %in% material)
    if (nrow(check_empty_frame)==0){
      intensity[nrow(intensity) + 1,] <- list(tech, material, 0)
    }
  }
}

#Estimate summary statistics of material intensity data
intensity_mean = intensity%>% group_by(technology,Material) %>% summarise(max = max(value),min = min(value),value = mean(value)) 
intensity_mean <- subset(intensity_mean, technology %in% repeat_technologies) #take the mean of intensity values to get one value per tech/material combo.

###ADD DISTRIBUTION FITTING CODE HERE###
```

**Material demand simulation (using mean material intensity values and Bistline et al. (2023) capacity):** This code simulates material demand using mean material intensity values and added capacities estimated based on Bistline et al. (2023).

```{r message=FALSE, warning=FALSE}
#Capacity dataframes subsetted for specific technologies and years
capacities_new <- subset(multi_model_capacities, technology  %in% repeat_technologies)
capacities_new<- subset(capacities_new, year %in% c(2025,2030,2035))#this was just to make this easier for key years 

##materials_demand: annual material demand by material, scenario, year, and technology 
#Material demand simulation
i=1
invisible(
for (material in materials){
  df_int = subset(intensity_mean, Material %in% material)
  df_cap = capacities_new %>% left_join (df_int, by='technology') 
  df_cap$totaldemand = df_cap$newval *df_cap$value
  colnames(df_cap)[colnames(df_cap) == "totaldemand"] = material
  df_cap = subset(df_cap, select = -c(Material, newval, value))
  
  if (i==1){
    materials_demand = df_cap
  }
  else{
    materials_demand <- materials_demand %>% left_join(df_cap, by=  c("scenario", "year", "technology"))
    
  }
  i= i+1
})
materials_demand <- materials_demand %>%
 select(-max.x, -min.x, -max.y, -min.y, -min, -max)

##materials_demand_sum: material demand by material, scenario, and year 
materials_demand_sum <-materials_demand %>% group_by(year,scenario) %>% summarise(Aluminum = sum(Aluminum), Cement = sum(Cement), Steel = sum(Steel))
                                                                                  #, Copper = sum(Cu), Silicon = sum(Si), Glass =sum(Glass), Indium = sum(In), Neodymium = sum(Nd), Silver= sum(Ag), Dysprosium  = sum(Dy), Gallium = sum(Ga), Selenium= sum(Se), Nickel = sum(Ni), Mangenese= sum(Mn))

```

**Material demand simulation (using mean material intensity values and Jenkins et al. (2023) capacity):** This code simulates material demand using mean material intensity values and added capacities estimated based on Jenkins et al. (2023).

```{r message=FALSE, warning=FALSE}
#Capacity dataframes subsetted for specific technologies and years
repeat_capacities_new <- subset(repeat_capacities_new, technology  %in% repeat_technologies)
repeat_capacities_new<- subset(repeat_capacities_new, year %in% c(2025,2030,2035))#this was just to make this easier
names(repeat_capacities_new)[names(repeat_capacities_new) == "run.name"] <- "scenario"
repeat_capacities_new$scenario[repeat_capacities_new$scenario == 'ira_mid'] <- 'IRA'
repeat_capacities_new$scenario[repeat_capacities_new$scenario == 'baseline'] <- 'REF'

##repeat_materials_demand: annual material demand by material, scenario, year, and technology 
#Material demand simulation
i=1
invisible(
for (material in materials){
  df_int = subset(intensity_mean, Material %in% material)
  df_cap = repeat_capacities_new %>% left_join (df_int, by='technology') 
  df_cap$totaldemand = df_cap$newval *df_cap$value
  colnames(df_cap)[colnames(df_cap) == "totaldemand"] = material
  df_cap = subset(df_cap, select = -c(Material, newval, value))
  
  if (i==1){
    repeat_materials_demand = df_cap
  }
  else{
    repeat_materials_demand <- repeat_materials_demand %>% left_join(df_cap, by=  c("scenario", "year", "technology"))
    
  }
  i= i+1
})
repeat_materials_demand <- repeat_materials_demand %>%
 select(-max.x, -min.x, -max.y, -min.y, -min, -max)

##repeat_materials_demand_sum: material demand by material, scenario, and year 
repeat_materials_demand_sum <-repeat_materials_demand %>% group_by(year,scenario) %>% summarise(Aluminum = sum(Aluminum), Cement = sum(Cement), Steel = sum(Steel))
#, Copper = sum(Cu), Silicon = sum(Si), Glass =sum(Glass), Indium = sum(In), Neodymium = sum(Nd), Silver= sum(Ag), Dysprosium  = sum(Dy), Gallium = sum(Ga), Selenium= sum(Se), Nickel = sum(Ni), Manganese= sum(Mn))

```

**Material intensity simulation (using material intensity distributions and Bistline et al. (2023) capacity):** This code simulates material demand using fitted material intensity distributions and added capacities estimated based on Bistline et al. (2023).

```{r}
#Material intensity simulation
set.seed(25)
n = 1 #number of iterations
df_mc <- data.frame() 

for (mat in materials){
  for (tech in repeat_technologies){
    loc = subset(intensity, Material %in% mat & technology %in% tech)
    if (nrow(loc)>1){
      minv = min(loc$value)
      maxv = max(loc$value)
      meanv = mean(loc$value)
      index_name = paste(mat, gsub("[[:space:]]", ".", tech), sep = ".")#addressing issues with spaces and dashes
      index_name = gsub("-", ".", index_name)
      df2 = data.frame(index = index_name, min = minv, max= maxv, mean = meanv)
      
      if (n==1){
        mc_values = df2
      }
      else{
        mc_values = rbind(mc_values, df2)
      }
      result <- rtriangle(100, minv, maxv, meanv)
      
      if(n==1){
        df_mc <- data.frame(cbind(result))
      }
      else{
        df_mc <- data.frame(cbind(df_mc, result))
      }
      colnames(df_mc)[colnames(df_mc) == "result"] =index_name
      n= n+1
    }
  }
}

```

**Material demand simulation (using material intensity distributions and Jenkins et al. (2023) capacity):** This code simulates material demand using fitted material intensity distributions and added capacities estimated based on Jenkins et al. (2023).

```{r}

# updated material inputs
#mulitplying material intensity from df_mc (run 10000 times) by the new capacities 
#result mc_material_demand is new material inputs needed per year. 

j=1
n=1

for (tech in repeat_technologies){
  for (mat in materials){
    for (scen in scenarios){
      n=1
      entry_df = data.frame()
      name = paste(mat, tech, sep ='.')
      name = str_replace_all(name, "-", ".")
      name = str_replace_all(name, ' ', '.')#addressing issues with spaces and dashes
      
      for (setyear in c(unique(capacities_new$year))){
        tryCatch(
          expr = {
            mc_loc = subset(df_mc, select = c(name))
            colnames(mc_loc)[colnames(mc_loc) == name] ="value"
            mc_loc = cbind(mc_loc, Material = mat, technology = tech, scenario = scen )
            cap_loc = subset(capacities_new, scenario %in% scen & technology %in% tech & year %in% setyear)
            test = merge(mc_loc, cap_loc)
            
            if (nrow(test)<1){#if there is no data setting value to -1
              test <- data.frame(matrix(-1,nrow = 100,ncol = 1))
              colnames(test)<- paste("a", setyear, sep = "")
              test$scenario = scen
              test$technology = tech
              test$value = -1 
              test$newval = -1
              test$Material = mat
              test$year = paste("a", setyear, sep = "")
            }
            
            else{
              test$value = as.numeric(test$value)
              test$finalvalue = test$value *test$newval
              colnames(test)[colnames(test) == "finalvalue"] = paste("a", setyear, sep = "")
            }
            
            if ((n==1) & (nrow(test)>1)){
              entry_df = test
              n = n+1
            }
            
            else if ((n!=1) & (nrow(test)>1)){
              df = subset(test, select = -c(Material, scenario,technology, newval,year,value))
              entry_df = cbind(entry_df, df)
            }
          },
          error = function(e){
          }
        )
      }
      
      if (j==1){
        mc_material_demand=entry_df
        j= j+1
      }
      else{
        mc_material_demand = rbind(mc_material_demand, entry_df)
      }
    } 
  }
}
```

```{r}
#same thing as above with repeat
j=1
n=1

for (tech in repeat_technologies){
  for (mat in materials){
    for (scen in "IRA"){
      n=1
      entry_df = data.frame()
      name = paste(mat, tech, sep ='.')
      name = str_replace_all(name, "-", ".")
      name = str_replace_all(name, ' ', '.')#addressing issues with spaces and dashes
      
      for (setyear in c(unique(repeat_capacities_new$year))){
        tryCatch(
          expr = {
            mc_loc = subset(df_mc, select = c(name))
            colnames(mc_loc)[colnames(mc_loc) == name] ="value"
            mc_loc = cbind(mc_loc, Material = mat, technology = tech, scenario = scen )
            cap_loc = subset(repeat_capacities_new, scenario %in% scen & technology %in% tech & year %in% setyear)
            test = merge(mc_loc, cap_loc)
            
            if (nrow(test)<1){#if there is no data setting value to -1
              test <- data.frame(matrix(-1,nrow = 100,ncol = 1))
              colnames(test)<- paste("a", setyear, sep = "")
              test$scenario = scen
              test$technology = tech
              test$value = -1 
              test$newval = -1
              test$Material = mat
              test$year = paste("a", setyear, sep = "")
            }
            
            else{
              test$value = as.numeric(test$value)
              test$finalvalue = test$value *test$newval
              colnames(test)[colnames(test) == "finalvalue"] = paste("a", setyear, sep = "")
            }
            
            if ((n==1) & (nrow(test)>1)){
              entry_df = test
              n = n+1
            }
            
            else if ((n!=1) & (nrow(test)>1)){
              df = subset(test, select = -c(Material, scenario,technology, newval,year,value))
              entry_df = cbind(entry_df, df)
            }
          },
          error = function(e){
          }
        )
      }
      
      if (j==1){
        repeat_mc_material_demand=entry_df
        j= j+1
      }
      else{
        repeat_mc_material_demand = rbind(repeat_mc_material_demand, entry_df)
      }
    } 
  }
}
```

```{r}
#sensitivity analysis plots 
par(mfrow = c(3, 3))  
par(mar = c(2, 2, 2, 2)) 

for(material in materials){
  for(tech in repeat_technologies){
    tryCatch(                      
      expr = {                      
        type = paste(material, tech, sep = ".")
        type = str_replace_all(type, " ", ".")
        type = str_replace_all(type, "-", ".")
        
        histtype = str_replace_all(type, "-", ".")
        histtype = str_replace_all(histtype, " ", ".")
        
        axislabels= subset(mc_values, index %in% type)
        hist_df = subset(df_mc, select = c(histtype))
        
        colnames(hist_df)[colnames(hist_df) == histtype] ="value"
        
        hist(as.numeric(hist_df$value), col = "antiquewhite1", cex.main=1.4,xaxt ='n', xlab = '', main = type, xlim =c(axislabels$min,axislabels$max))
        axis(1, at = c(axislabels$mean, axislabels$min,axislabels$max),
             labels = c(round(axislabels$mean, digits =5), round(axislabels$min, digits = 5), round(axislabels$max, digits = 5)))
      },
      error = function(e){
        dummy = '1'
      }
    )
  }
}
```

```{r}

plot.new()

for(scen in scenarios){
  
  par(mfrow = c(3, 3)) 
  par(mar=c(2,2,2,2)) 
  par(xpd=NA)
  

  for(mat in materials){
    for(tech in repeat_technologies){
      tryCatch(                       
        
        expr = { 
          boxplot_df = subset(mc_material_demand, technology %in% tech & scenario %in% scen & Material %in% mat)
          boxplot_df =subset(boxplot_df, select = c(a2025, a2030, a2035))
          if (nrow(boxplot_df)>0){
              boxplot(boxplot_df, cex.main=1.2,cex.axis=1.2, main =  (paste (mat, tech, sep = ";")), names = c('2025','2030','2035'),border = "coral3", col ="white")
            
          }
        },
        error = function(e){
          message("error")
          }
      )
    }
  }
  
  title(main = scen, xlab= "test", outer=TRUE, cex.main = 1)

}

```

**Saving data frames to "outputs" folder:** This code saves the dataframes for use in the model 

```{r}
folder_path <- "C:/Users/avery/OneDrive/Desktop/MaterialDemand/Outputs"

file_path1 <- file.path(folder_path, "multi_model_materials_demand_sum.csv")
file_path2 <- file.path(folder_path, "repeat_materials_demand_sum.csv")
file_path3 <- file.path(folder_path, "multi_model_mc_material_demand.csv")
file_path4 <- file.path(folder_path, "repeat_mc_material_demand.csv")

# Write each dataframe to its respective CSV file in the specified folder
write.csv(materials_demand_sum, file = file_path1, row.names = FALSE)
write.csv(repeat_materials_demand_sum, file = file_path2, row.names = FALSE)
write.csv(mc_material_demand, file = file_path3, row.names = FALSE)
write.csv(repeat_mc_material_demand, file = file_path4, row.names = FALSE)

```
```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
