---
title: "Material Demand Simulation"
output:
  html_document:
    df_print: paged
date created: 11/28/23
date last modified: 12/13/23
created by: Avery Moorhead
last revised by: Erin Mayfield
---
```{r setup, include=FALSE}


# Set working directory using here::set_here()
knitr::opts_knit$set(root.dir = "C:/Users/avery/OneDrive/Desktop/MaterialDemand/Raw Data")

# Clear objects from workspace
rm(list = ls())

```

```{r message=FALSE, warning=FALSE}
###NTS: Currently only running this script on aluminum cement and steel. you can change this by changing rows 222,305, and 341
#Clear objects from workspace
rm(list = ls())

#Set working directory
#setwd("C:/Users/avery/OneDrive/Desktop/MaterialDemand")

#Load libraries
library(readxl)
library(data.table)
library(dplyr)
library(tidyverse)
library(triangle)
library(stringr)
library(knitr)
```

**Capacity inputs (Bistline et al., 2023):** This code processes electric power capacity input data from Bistline et al. (2023). The data include total capacity and added capacity by technology, capacity expansion model, and year. All capacity data are in units of gigawatts.

```{r message=FALSE, warning=FALSE}
##Cap_IRA: total capacity for each technology, model, and period under IRA scenario (units: gigawatts)

#Load data
Cap_IRA = read_excel("Multi Model Capacity Data.xlsx", 
                      sheet = "Cap IRA", na="ND", range =  cell_rows(1:18), col_names= FALSE)

#Creating dataframe and processing data
Cap_IRA <- t(Cap_IRA)
Cap_IRA <- as.data.frame(Cap_IRA)
rownames(Cap_IRA)<-NULL
new_colnames <- unlist(Cap_IRA[1, ])
colnames(Cap_IRA) <- new_colnames
Cap_IRA <- Cap_IRA[-1,]
colnames(Cap_IRA)[1] ="model"
colnames(Cap_IRA)[2] ="year"
Cap_IRA$year[is.na(Cap_IRA$year)] <- "2021"
Cap_IRA <-  Cap_IRA[(1:34),]
Cap_IRA[is.na(Cap_IRA)] <- 0
Cap_IRA[] <- lapply(Cap_IRA, gsub, pattern = "N/A", replacement = "0", fixed = TRUE)
Cap_IRA <- as.data.frame(Cap_IRA)

#Creating dataframe for total capacity in the year 2021
initial_year <- subset(Cap_IRA, year %in% "2021")
initial_year[is.na(initial_year)] <- '0'

#Additional data processing 
Cap_IRA = Cap_IRA[-1,]
Cap_IRA[is.na(Cap_IRA)] <- '0'
Cap_IRA <- rbind(Cap_IRA, initial_year)
Cap_IRA <- reshape2::melt(Cap_IRA, id = c("model","year")) 
Cap_IRA$value <- as.numeric(unlist(Cap_IRA$value))
Cap_IRA$scenario <- "IRA"

##Y_Cap_IRA: additional capacity for each technology, model, and period under IRA scenario (units: gigawatts)

#Load data
Y_Cap_IRA = read_excel("Multi Model Capacity Data.xlsx", 
                     sheet = "Cap IRA", na="ND", col_names= FALSE, range = cell_rows(28:45))

#Creating dataframe and processing data
Y_Cap_IRA <- t(Y_Cap_IRA)
Y_Cap_IRA <- as.data.frame(Y_Cap_IRA)
rownames(Y_Cap_IRA)<-NULL
new_colnames <- unlist(Y_Cap_IRA[1, ])
colnames(Y_Cap_IRA) <- new_colnames
Y_Cap_IRA <- Y_Cap_IRA[-1,]
colnames(Y_Cap_IRA)[1] ="model"
colnames(Y_Cap_IRA)[2] ="year"
Y_Cap_IRA$year[is.na(Y_Cap_IRA$year)] <- "2021"
Y_Cap_IRA <-  Y_Cap_IRA[(1:34),]
Y_Cap_IRA[is.na(Y_Cap_IRA)] <- 0
Y_Cap_IRA[] <- lapply(Y_Cap_IRA, gsub, pattern = "N/A", replacement = "0", fixed = TRUE)
Y_Cap_IRA <- as.data.frame(Y_Cap_IRA)
Y_Cap_IRA[is.na(Y_Cap_IRA)] <- '0'
Y_Cap_IRA <- reshape2::melt(Y_Cap_IRA, id = c("model","year")) 
Y_Cap_IRA$value <- as.numeric(unlist(Y_Cap_IRA$value))
Y_Cap_IRA$scenario <- "IRA"
Y_Cap_IRA$model[Y_Cap_IRA$model=='0']<-"2021"

##Cap_Ref: total capacity for each technology, model, and period under reference scenario (units: gigawatts)

#Load data
Cap_Ref = read_excel("Multi Model Capacity Data.xlsx", 
                     sheet = "Cap Ref", na="ND", range =  cell_rows(1:18), col_names= FALSE)

#Creating dataframe and processing data
Cap_Ref <- t(Cap_Ref)
Cap_Ref <- as.data.frame(Cap_Ref)
rownames(Cap_Ref)<-NULL
new_colnames <- unlist(Cap_Ref[1, ])
colnames(Cap_Ref) <- new_colnames
Cap_Ref <- Cap_Ref[-1,]
colnames(Cap_Ref)[1] ="model"
colnames(Cap_Ref)[2] ="year"
Cap_Ref$year[is.na(Cap_Ref$year)] <- "2021"
Cap_Ref <-  Cap_Ref[(1:34),]
Cap_Ref[is.na(Cap_Ref)] <- 0
Cap_Ref[] <- lapply(Cap_Ref, gsub, pattern = "N/A", replacement = "0", fixed = TRUE)
Cap_Ref <- as.data.frame(Cap_Ref)
Cap_Ref = Cap_Ref[-1,]
Cap_Ref[is.na(Cap_Ref)] <- '0'
Cap_Ref <- rbind(Cap_Ref, initial_year)
Cap_Ref <- reshape2::melt(Cap_Ref, id = c("model","year")) 
Cap_Ref$value <- as.numeric(unlist(Cap_Ref$value))
Cap_Ref$scenario <- "Ref"

##Y_Cap_Ref: additional capacity for each technology, model, and period under reference scenario (units: gigawatts)

#Load data
Y_Cap_Ref = read_excel("Multi Model Capacity Data.xlsx", 
                       sheet = "Cap Ref", na="ND", col_names= FALSE, range = cell_rows(28:45))

#Creating dataframe and processing data
Y_Cap_Ref <- t(Y_Cap_Ref)
Y_Cap_Ref <- as.data.frame(Y_Cap_Ref)
rownames(Y_Cap_Ref)<-NULL
new_colnames <- unlist(Y_Cap_Ref[1, ])
colnames(Y_Cap_Ref) <- new_colnames
Y_Cap_Ref <- Y_Cap_Ref[-1,]
colnames(Y_Cap_Ref)[1] ="model"
colnames(Y_Cap_Ref)[2] ="year"
Y_Cap_Ref$year[is.na(Y_Cap_Ref$year)] <- "2021"
Y_Cap_Ref <-  Y_Cap_Ref[(1:34),]
Y_Cap_Ref[is.na(Y_Cap_Ref)] <- 0
Y_Cap_Ref[] <- lapply(Y_Cap_Ref, gsub, pattern = "N/A", replacement = "0", fixed = TRUE)
Y_Cap_Ref <- as.data.frame(Y_Cap_Ref)
Y_Cap_Ref[is.na(Y_Cap_Ref)] <- '0'
Y_Cap_Ref <- reshape2::melt(Y_Cap_Ref, id = c("model","year")) 
Y_Cap_Ref$value <- as.numeric(unlist(Y_Cap_Ref$value))
Y_Cap_Ref$scenario <- "Ref"
Y_Cap_Ref$model[Y_Cap_Ref$model=='0']<-"2021"

##capacities_all: total capacity for each technology, model, and period under reference and IRA scenarios (units: gigawatts)

#Merging Cap_IRA and Cap_Ref dataframes and data processing
capacities_all <- rbind(Cap_IRA,Cap_Ref)
colnames(capacities_all)[3] ="technology"
capacities_all$technology <- gsub("Wind", "onshore wind", capacities_all$technology)
capacities_all$technology <- gsub("Solar Utility", "utility-scale solar pv", capacities_all$technology)

##capacities: mean, minimum, and maximum total capacity across models and for each technology and period under reference and IRA scenarios (units: gigawatts)
capacities <- capacities_all[-1]
capacities <-capacities %>% group_by(scenario,year,technology) %>% summarise(newval = mean(value), max = max(value),min = min(value)) #calculating mean, minimum, and maximum capacities across models 

##capacities_y_all: annualized additional capacity for each technology, model, and period under reference and IRA scenarios (units: gigawatts)

#Merging Y_Cap_IRA and Y_Cap_Ref dataframes
capacities_y_all <- rbind(Y_Cap_IRA,Y_Cap_Ref)

#Annualizing capacity data 
capacities_y_all$newval <- capacities_y_all$value
capacities_y_all$newval[capacities_y_all$year == 2025] <- capacities_y_all$value[capacities_y_all$year == 2025] / 4
capacities_y_all$newval[capacities_y_all$year == 2030] <- capacities_y_all$value[capacities_y_all$year == 2030] / 5
capacities_y_all$newval[capacities_y_all$year == 2035] <- capacities_y_all$value[capacities_y_all$year == 2035] / 5

##capacities_y: mean annualized additional capacity across models and for each technology and period under reference and IRA scenarios (units: gigawatts)
colnames(capacities_y_all)[3] ="technology"
capacities_y_all$technology <- gsub("Wind", "onshore wind", capacities_y_all$technology)
capacities_y <- capacities_y_all[c(-1,-4)]
capacities_y <-   capacities_y %>% group_by(scenario,year,technology) %>% summarise(newval = mean(newval)) #calculating mean capacities across models 

##final_y_cap_IRA: additional capacity for each technology, model and year under reference and IRA scenarios (units: gigawatts)

#Creating intermediate dataframes for each period and data processing
temp_df_2025 <- capacities_y %>%
  filter(year == 2025)
duplicated_vector <- rep(1:nrow(temp_df_2025), each = 4)
temp_df_2025 <- temp_df_2025[duplicated_vector, ]

temp_df_2030 <- capacities_y %>%
  filter(year == 2030)
duplicated_vector <- rep(1:nrow(temp_df_2030), each = 5)
temp_df_2030 <- temp_df_2030[duplicated_vector, ]
num_rows <- nrow(temp_df_2030)

temp_df_2035 <- capacities_y %>%
  filter(year == 2035)
duplicated_vector <- rep(1:nrow(temp_df_2035), each = 5)
temp_df_2035 <- temp_df_2035[duplicated_vector, ]
num_rows <- nrow(temp_df_2035)

temp_df_2025 <- temp_df_2025 %>% group_by(technology,scenario)%>%
  mutate(year = ifelse(row_number() %% 3 == 0, 2022, year)) %>% mutate(year = ifelse((row_number() - 1) %% 3 == 0 & row_number() != 1, 2023, year))%>% mutate(year = ifelse((row_number() - 2) %% 3 == 0 & row_number() != 1, 2024, year))

temp_df_2030 <- temp_df_2030 %>% group_by(technology,scenario)%>%
  mutate(year = ifelse(row_number() %% 4 == 0, 2029, year)) %>% mutate(year = ifelse((row_number() - 1) %% 4 == 0 & row_number() != 1, 2028, year))%>% mutate(year = ifelse((row_number() - 2) %% 4 == 0 & row_number() != 1, 2027, year))%>% mutate(year = ifelse((row_number() - 3) %% 4 == 0 & row_number() != 1, 2026, year))

temp_df_2035 <- temp_df_2035 %>% group_by(technology,scenario)%>%
  mutate(year = ifelse(row_number() %% 4 == 0, 2034, year)) %>% mutate(year = ifelse((row_number() - 1) %% 4 == 0 & row_number() != 1, 2033, year))%>% mutate(year = ifelse((row_number() - 2) %% 4 == 0 & row_number() != 1, 2032, year))%>% mutate(year = ifelse((row_number() - 3) %% 4 == 0 & row_number() != 1, 2031, year))

#Combining intermediate dataframes for each period 
final_y_cap_IRA <- rbind(temp_df_2030,temp_df_2035,temp_df_2025)
final_y_cap_IRA$technology <- ifelse(final_y_cap_IRA$technology == "wind", "onshore wind", final_y_cap_IRA$technology)
final_y_cap_IRA["technology"][final_y_cap_IRA["technology"] == "Solar Utility"] <- "utility-scale solar pv"###NTS: check solar distributed vs utility 

```

**Capacity inputs (Jenkins et al., 2023):** This code uploads and processes electric power capacity input data from Jenkins et al. (2023). The data include total capacity and added capacity by technology, capacity expansion model, and year. All capacity data are in units of gigawatts.

```{r message=FALSE, warning=FALSE}
##repeat_capacities: total capacity for each technology and year under reference and IRA scenarios (units: gigawatts)

#Load data
input_capacities= read.csv("capacity_y.csv")

#Creating dataframe and processing data
scenarios = c("Ref", "IRA")#Define the REPEAT scenarios to be included
repeat_technologies = c("onshore wind", "utility-scale solar pv", "Nuclear New", "Wind Offshore")#Define the REPEAT technologies to be included
repeat_scenarios = c("baseline", "ira_mid")
input_capacities["tech..outputs_group_detailed"][input_capacities["tech..outputs_group_detailed"] == "offshore wind"] <- "offshore wind fixed"
input_capacities["tech..outputs_group_detailed"][input_capacities["tech..outputs_group_detailed"] == "new nuclear"] <- "Nuclear New"
repeat_capacities = subset(input_capacities, run.name %in% repeat_scenarios & tech..outputs_group_detailed %in% repeat_technologies) %>% group_by(run.name,year,tech..outputs_group_detailed) %>% summarise(newval = sum(value))

##repeat_capacities_new: annualized added capacity for each technology and year under reference and IRA scenarios (units: gigawatts)

#Load data
input_capacities_delta = read.csv("capacity_delta_y.csv")

#Creating dataframe and processing data
input_capacities_delta["tech..outputs_group_detailed"][input_capacities_delta["tech..outputs_group_detailed"] == "offshore wind"] <- "offshore wind fixed"
input_capacities_delta["tech..outputs_group_detailed"][input_capacities_delta["tech..outputs_group_detailed"] == "new nuclear"] <- "Nuclear New"
repeat_capacities_new = subset(input_capacities_delta, run.name %in% repeat_scenarios & type == "new" & tech..outputs_group_detailed %in% repeat_technologies) %>%
  group_by(run.name,year,tech..outputs_group_detailed) %>% summarise(newval = sum(value))
colnames(repeat_capacities_new)[colnames(repeat_capacities_new) == "tech..outputs_group_detailed"] ="technology"

#Annualizing added capacities
repeat_capacities_new<- subset(repeat_capacities_new, year %in% c("2026","2030","2035"))
repeat_capacities_new$newval <- ifelse(repeat_capacities_new$year == 2026, repeat_capacities_new$newval / 2, repeat_capacities_new$newval)
repeat_capacities_new$year <- ifelse(repeat_capacities_new$year == 2026, 2025, repeat_capacities_new$year)

```

**Historical material supply, demand, imports, and exports:** This code uploads historical U.S. consumption, production, imports, and exports of materials for 2018 to 2022 from the U.S. Geological Survey (2023).

```{r message=FALSE, warning=FALSE}
##inputs_usgs: U.S. supply, demand, imports, and exports for materials from 2018 to 2021

#Load data
input_usgs = read_excel("US Geological Survey 2023/Data by material/USGS_2023_IRA materials calculations.xlsx", 
                        sheet = "Combined", range = cell_cols(1:5))
#Filtering of USGS data
usgs_variables = c("Consumption - US", "Production - US", "Import - US", "Export - US")
flow = subset(input_usgs, Variable %in%usgs_variables)
```

**Material intensity (Liang et al., 202X; Wang et al., 202X):** This code uploads and processes material intensity values from Liang et al. (202X) and Wang et al. (202X).

```{r message=FALSE, warning=FALSE}
##repeat_capacities_delta: annualized added capacity for each technology and year under reference and IRA scenarios (units: gigawatts)

#Load data
input_wang = read_excel("Final Intensity Values.xlsx", sheet = "Intensity Values", range = cell_cols(1:6))

#Creating dataframe and processing data
colnames(input_wang)[1] ="technology"
colnames(input_wang)[3] ="value"
input_wang <- input_wang[!duplicated(input_wang), ]
input_wang <- input_wang[input_wang$`Year of the data` <= 2025, 
                                  !(names(input_wang) %in% c('Year of the data', 'Authors', 'Publish year'))]
###NTS:data processing, worth checking some of these 
input_wang["technology"][input_wang["technology"] == "Nuclear"] <- "Nuclear New"
input_wang["technology"][input_wang["technology"] == "Onshore_AG"] <- "onshore wind"
input_wang["technology"][input_wang["technology"] == "Onshore_DD_PMG"] <- "onshore wind"
input_wang["technology"][input_wang["technology"] == "Wind Directed_Drive"] <- "onshore wind"
input_wang["technology"][input_wang["technology"] == "Geared"] <- "onshore wind"
input_wang["technology"][input_wang["technology"] == "Wind onshore"] <- "onshore wind"
input_wang["technology"][input_wang["technology"] == "Wind Onshore"] <- "onshore wind"
input_wang["technology"][input_wang["technology"] == "CSI_PV"] <- "utility-scale solar pv"
input_wang$technology[input_wang$technology == "Offshore_DD_PMG"] <- "Wind Offshore"
input_wang$technology[input_wang$technology == "Offshore_AG"] <- "Wind Offshore"
input_wang$Material[input_wang$Material == "Neodymium "] <-  "Neodymium"

```

**Material intensity summary statistics and distribution fitting:** This code estimates summary statistics and fits probability distributions to material intensity data from Liang et al. (202X) and Wang et al. (202X).

```{r message=FALSE, warning=FALSE}
#Subset material intensity data to include some materials 
materials <- c("Aluminum","Cement","Steel")
#print(materials)# Print the unique materials
intensity = subset(input_wang, Material %in% materials)

#Check if all material/technology scenarios are present, if not add 0.
for (material in materials) {
  for (tech in repeat_technologies){  
    check_empty_frame = subset(intensity, technology %in% tech & Material %in% material)
    if (nrow(check_empty_frame)==0){
      intensity[nrow(intensity) + 1,] <- list(tech, material, 0)
    }
  }
}

#Estimate summary statistics of material intensity data
intensity_mean = intensity%>% group_by(technology,Material) %>% summarise(max = max(value),min = min(value),value = mean(value)) 
intensity_mean <- subset(intensity_mean, technology %in% repeat_technologies) #take the mean of intensity values to get one value per tech/material combo.

###ADD DISTRIBUTION FITTING CODE HERE###
```

**Material demand simulation (using mean material intensity values and Bistline et al. (2023) capacity):** This code simulates material demand using mean material intensity values and added capacities estimated based on Bistline et al. (2023).

```{r message=FALSE, warning=FALSE}
#Capacity dataframes subsetted for specific technologies and years
capacities_new <- subset(final_y_cap_IRA, technology  %in% repeat_technologies)
capacities_new<- subset(capacities_new, year %in% c(2025,2030,2035))#this was just to make this easier for key years 

##materials_demand: annual material demand by material, scenario, year, and technology 
#Material demand simulation
i=1
invisible(
for (material in materials){
  df_int = subset(intensity_mean, Material %in% material)
  df_cap = capacities_new %>% left_join (df_int, by='technology') 
  df_cap$totaldemand = df_cap$newval *df_cap$value
  colnames(df_cap)[colnames(df_cap) == "totaldemand"] = material
  df_cap = subset(df_cap, select = -c(Material, newval, value))
  
  if (i==1){
    materials_demand = df_cap
  }
  else{
    materials_demand <- materials_demand %>% left_join(df_cap, by=  c("scenario", "year", "technology"))
    
  }
  i= i+1
})
materials_demand <- materials_demand %>%
 select(-max.x, -min.x, -max.y, -min.y, -min, -max)

##materials_demand_sum: material demand by material, scenario, and year 
materials_demand_sum <-materials_demand %>% group_by(year,scenario) %>% summarise(Aluminum = sum(Aluminum), Cement = sum(Cement), Steel = sum(Steel))
                                                                                  #, Copper = sum(Cu), Silicon = sum(Si), Glass =sum(Glass), Indium = sum(In), Neodymium = sum(Nd), Silver= sum(Ag), Dysprosium  = sum(Dy), Gallium = sum(Ga), Selenium= sum(Se), Nickel = sum(Ni), Mangenese= sum(Mn))

```

**Material demand simulation (using mean material intensity values and Jenkins et al. (2023) capacity):** This code simulates material demand using mean material intensity values and added capacities estimated based on Jenkins et al. (2023).

```{r message=FALSE, warning=FALSE}
#Capacity dataframes subsetted for specific technologies and years
repeat_capacities_new <- subset(repeat_capacities_new, technology  %in% repeat_technologies)
repeat_capacities_new<- subset(repeat_capacities_new, year %in% c(2025,2030,2035))#this was just to make this easier
names(repeat_capacities_new)[names(repeat_capacities_new) == "run.name"] <- "scenario"
repeat_capacities_new$scenario[repeat_capacities_new$scenario == 'ira_mid'] <- 'IRA'
repeat_capacities_new$scenario[repeat_capacities_new$scenario == 'baseline'] <- 'REF'

##repeat_materials_demand: annual material demand by material, scenario, year, and technology 
#Material demand simulation
i=1
invisible(
for (material in materials){
  df_int = subset(intensity_mean, Material %in% material)
  df_cap = repeat_capacities_new %>% left_join (df_int, by='technology') 
  df_cap$totaldemand = df_cap$newval *df_cap$value
  colnames(df_cap)[colnames(df_cap) == "totaldemand"] = material
  df_cap = subset(df_cap, select = -c(Material, newval, value))
  
  if (i==1){
    repeat_materials_demand = df_cap
  }
  else{
    repeat_materials_demand <- repeat_materials_demand %>% left_join(df_cap, by=  c("scenario", "year", "technology"))
    
  }
  i= i+1
})
repeat_materials_demand <- repeat_materials_demand %>%
 select(-max.x, -min.x, -max.y, -min.y, -min, -max)

##repeat_materials_demand_sum: material demand by material, scenario, and year 
repeat_materials_demand_sum <-repeat_materials_demand %>% group_by(year,scenario) %>% summarise(Aluminum = sum(Aluminum), Cement = sum(Cement), Steel = sum(Steel))
#, Copper = sum(Cu), Silicon = sum(Si), Glass =sum(Glass), Indium = sum(In), Neodymium = sum(Nd), Silver= sum(Ag), Dysprosium  = sum(Dy), Gallium = sum(Ga), Selenium= sum(Se), Nickel = sum(Ni), Manganese= sum(Mn))

```

**Material intensity simulation (using material intensity distributions and Bistline et al. (2023) capacity):** This code simulates material demand using fitted material intensity distributions and added capacities estimated based on Bistline et al. (2023).

```{r}
#Material intensity simulation
set.seed(25)
n = 1 #number of iterations
df_mc <- data.frame() 

for (mat in materials){
  for (tech in repeat_technologies){
    loc = subset(intensity, Material %in% mat & technology %in% tech)
    if (nrow(loc)>1){
      minv = min(loc$value)
      maxv = max(loc$value)
      meanv = mean(loc$value)
      index_name = paste(mat, gsub("[[:space:]]", ".", tech), sep = ".")#addressing issues with spaces and dashes
      index_name = gsub("-", ".", index_name)
      df2 = data.frame(index = index_name, min = minv, max= maxv, mean = meanv)
      
      if (n==1){
        mc_values = df2
      }
      else{
        mc_values = rbind(mc_values, df2)
      }
      result <- rtriangle(100, minv, maxv, meanv)
      
      if(n==1){
        df_mc <- data.frame(cbind(result))
      }
      else{
        df_mc <- data.frame(cbind(df_mc, result))
      }
      colnames(df_mc)[colnames(df_mc) == "result"] =index_name
      n= n+1
    }
  }
}

```

**Material demand simulation (using material intensity distributions and Jenkins et al. (2023) capacity):** This code simulates material demand using fitted material intensity distributions and added capacities estimated based on Jenkins et al. (2023).

```{r}

# updated material inputs
#mulitplying material intensity from df_mc (run 10000 times) by the new capacities 
#result mc_material_demand is new material inputs needed per year. 

j=1
n=1

for (tech in repeat_technologies){
  for (mat in materials){
    for (scen in scenarios){
      n=1
      entry_df = data.frame()
      name = paste(mat, tech, sep ='.')
      name = str_replace_all(name, "-", ".")
      name = str_replace_all(name, ' ', '.')#addressing issues with spaces and dashes
      
      for (setyear in c(unique(capacities_new$year))){
        tryCatch(
          expr = {
            mc_loc = subset(df_mc, select = c(name))
            colnames(mc_loc)[colnames(mc_loc) == name] ="value"
            mc_loc = cbind(mc_loc, Material = mat, technology = tech, scenario = scen )
            cap_loc = subset(capacities_new, scenario %in% scen & technology %in% tech & year %in% setyear)
            test = merge(mc_loc, cap_loc)
            
            if (nrow(test)<1){#if there is no data setting value to -1
              test <- data.frame(matrix(-1,nrow = 100,ncol = 1))
              colnames(test)<- paste("a", setyear, sep = "")
              test$scenario = scen
              test$technology = tech
              test$value = -1 
              test$newval = -1
              test$Material = mat
              test$year = paste("a", setyear, sep = "")
            }
            
            else{
              test$value = as.numeric(test$value)
              test$finalvalue = test$value *test$newval
              colnames(test)[colnames(test) == "finalvalue"] = paste("a", setyear, sep = "")
            }
            
            if ((n==1) & (nrow(test)>1)){
              entry_df = test
              n = n+1
            }
            
            else if ((n!=1) & (nrow(test)>1)){
              df = subset(test, select = -c(Material, scenario,technology, newval,year,value))
              entry_df = cbind(entry_df, df)
            }
          },
          error = function(e){
          }
        )
      }
      
      if (j==1){
        mc_material_demand=entry_df
        j= j+1
      }
      else{
        mc_material_demand = rbind(mc_material_demand, entry_df)
      }
    } 
  }
}
```

```{r}
#same thing as above with repeat
j=1
n=1

for (tech in repeat_technologies){
  for (mat in materials){
    for (scen in "IRA"){
      n=1
      entry_df = data.frame()
      name = paste(mat, tech, sep ='.')
      name = str_replace_all(name, "-", ".")
      name = str_replace_all(name, ' ', '.')#addressing issues with spaces and dashes
      
      for (setyear in c(unique(repeat_capacities_new$year))){
        tryCatch(
          expr = {
            mc_loc = subset(df_mc, select = c(name))
            colnames(mc_loc)[colnames(mc_loc) == name] ="value"
            mc_loc = cbind(mc_loc, Material = mat, technology = tech, scenario = scen )
            cap_loc = subset(repeat_capacities_new, scenario %in% scen & technology %in% tech & year %in% setyear)
            test = merge(mc_loc, cap_loc)
            
            if (nrow(test)<1){#if there is no data setting value to -1
              test <- data.frame(matrix(-1,nrow = 100,ncol = 1))
              colnames(test)<- paste("a", setyear, sep = "")
              test$scenario = scen
              test$technology = tech
              test$value = -1 
              test$newval = -1
              test$Material = mat
              test$year = paste("a", setyear, sep = "")
            }
            
            else{
              test$value = as.numeric(test$value)
              test$finalvalue = test$value *test$newval
              colnames(test)[colnames(test) == "finalvalue"] = paste("a", setyear, sep = "")
            }
            
            if ((n==1) & (nrow(test)>1)){
              entry_df = test
              n = n+1
            }
            
            else if ((n!=1) & (nrow(test)>1)){
              df = subset(test, select = -c(Material, scenario,technology, newval,year,value))
              entry_df = cbind(entry_df, df)
            }
          },
          error = function(e){
          }
        )
      }
      
      if (j==1){
        repeat_mc_material_demand=entry_df
        j= j+1
      }
      else{
        repeat_mc_material_demand = rbind(repeat_mc_material_demand, entry_df)
      }
    } 
  }
}
```

```{r}
#sensitivity analysis plots 
par(mfrow = c(3, 3))  
par(mar = c(2, 2, 2, 2)) 

for(material in materials){
  for(tech in repeat_technologies){
    tryCatch(                      
      expr = {                      
        type = paste(material, tech, sep = ".")
        type = str_replace_all(type, " ", ".")
        type = str_replace_all(type, "-", ".")
        
        histtype = str_replace_all(type, "-", ".")
        histtype = str_replace_all(histtype, " ", ".")
        
        axislabels= subset(mc_values, index %in% type)
        hist_df = subset(df_mc, select = c(histtype))
        
        colnames(hist_df)[colnames(hist_df) == histtype] ="value"
        
        hist(as.numeric(hist_df$value), col = "antiquewhite1", cex.main=1.4,xaxt ='n', xlab = '', main = type, xlim =c(axislabels$min,axislabels$max))
        axis(1, at = c(axislabels$mean, axislabels$min,axislabels$max),
             labels = c(round(axislabels$mean, digits =5), round(axislabels$min, digits = 5), round(axislabels$max, digits = 5)))
      },
      error = function(e){
        dummy = '1'
      }
    )
  }
}
```

```{r}

plot.new()

for(scen in scenarios){
  
  par(mfrow = c(3, 3)) 
  par(mar=c(2,2,2,2)) 
  par(xpd=NA)
  

  for(mat in materials){
    for(tech in repeat_technologies){
      tryCatch(                       
        
        expr = { 
          boxplot_df = subset(mc_material_demand, technology %in% tech & scenario %in% scen & Material %in% mat)
          boxplot_df =subset(boxplot_df, select = c(a2025, a2030, a2035))
          if (nrow(boxplot_df)>0){
              boxplot(boxplot_df, cex.main=1.2,cex.axis=1.2, main =  (paste (mat, tech, sep = ";")), names = c('2025','2030','2035'),border = "coral3", col ="white")
            
          }
        },
        error = function(e){
          message("error")
          }
      )
    }
  }
  
  title(main = scen, xlab= "test", outer=TRUE, cex.main = 1)

}

```

**Saving data frames to "inputs" folder:** This code saves the dataframes for use in the model 



```

